---
title: "Equine Cryptorchid Stats"
author: "Gregg Griffenhagen"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
  #setup necessary packages and read in data
library(ggplot2)
library(magrittr)
library(tidyverse)
library(nnet)
library(emmeans)
library(forestplot)

data <- read.csv(file="crypt_data_2.csv", colClasses = c("NULL",
                                                       "factor",
                                                       "factor",
                                                       "NULL"))
```

## Evaluate the data graphically
```{r}
p1 <- ggplot(data=data, aes(brd, fill=lat))
p1 + geom_bar(position="fill")

p2 <- ggplot(data=data, aes(lat, fill=brd))
p2 + geom_bar(position="stack") +
   theme_classic() +
  labs(x="Laterality", y="Total Cases")
```

## Create contingency table to look at total numbers and evaluate whether there
## are any differences between groups

```{r}
     #create contingency table
cont_table <- table(data$brd, data$lat)
cont_table

     #are there any significant differences?
chisq.test(cont_table)
fisher.test(cont_table, simulate.p.value = TRUE)
```

## There are statistical differences in the total numbers between groups
## To evaluate differences between groups, a multinomial model was used and all
## pairs compared at each level - bilateral, right sided, left sided

```{r}
t1 <- as.data.frame.table(cont_table)
colnames(t1) <- c("brd", "lat", "freq")

t1$brd <- fct_relevel(t1$brd, "QH")  #relevel to compare all to QH

model2 <- multinom(lat ~ brd, data = t1, weights = freq, Hess=T, model=T)
summary(model2)

emmout <- emmeans(model2, "brd", by="lat")
contrast(emmout, "trt.vs.ctrl")
```
## There are no differences in the occurence of bilateral cryptorchidism
## There are differences in L v R for Arabians and TB vs QH
## Remove bilateral from model and repeat

```{r}
t1$lat <- fct_relevel(t1$lat, "Rt")
t2 <- t1 %>% filter(lat != "Bilat", .preserve=FALSE) %>%
  droplevels()

model3 <- multinom(lat ~ brd, data = t2, weights = freq)                      
summary(model3)

emmout2 <- emmeans(model3, "brd", by="lat")
contrast(emmout2, "trt.vs.ctrl")
```

## Same results, probably more relevant model
## Print odds ratios and confidence intervals
# These are the odds of a unilateral crypt being left sided (vs QH)

```{r}
odds.ratios <- enframe(exp(coef(model3)))
colnames(odds.ratios) <- c("breed", "odds.ratio")

confint <- as_tibble(exp(confint(model3)))
colnames(confint) <- c("lower_CI", "upper_CI")

OR_summary <- as_tibble(c(odds.ratios[-1,], confint[-1,]))
OR_summary
```

## Forestplot of the data - this is the odds ratio for a unilateral crypt horse
## to be left-sided vs. a Quarterhorse. Arabians and Thoroughbreds (and likely Ponies) are
## significantly less likely to present as left-sided than Quarterhorses - or those 3 are
## significantly more likely to present as right sided than QH.

```{r}
fp_data <- tibble(mean = OR_summary$odds.ratio,
                     lower = OR_summary$lower_CI,
                     upper = OR_summary$upper_CI,
                     breed = c("Arabian", "Draft", "Mixed", "Pony", "Saddlebred",
                               "Thoroughbred", "Warmblood"),
                     OR = round(OR_summary$odds.ratio, 3))
fp_data %>%
  forestplot(labeltext = c(breed, OR),
             clip = c(0, 8),
             vertices = TRUE,
             zero = 1,
             xlab = "Odds Ratio (vs. Quarterhorse)") %>%
  fp_append_row(breed = "Quarterhorse",
                OR = "Referrent",
                position = 1) %>%
  fp_add_header(breed = "Breed",
                OR = "OR") %>%
  fp_add_lines(h_2 = gpar(lty = 1), h_3 = gpar(columns = 1:2)) %>%
  fp_set_zebra_style("#f9f9f9")
```

